Improving legitimacy and efficiency \\ of decisions in democratic communities

### 2017-09-19 statistical-quorum.tex
Adding TODOs.
Front page.

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Improving legitimacy and efficiency of decisions in democratic communities}
%\def\plainauthor{First Author, Second Author, Third Author,
%  Fourth Author, Fifth Author, Sixth Author}
%\def\emptyauthor{}
%\def\plainkeywords{Authors' choice; of terms; separated; by
%  semicolons; include commas, within terms only; required.}
%\def\plaingeneralterms{Documentation, Standardization}

@@

\begin{abstract}
%The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
%While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
%different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
%maximize productivity?
%We propose a novel "statistical quorum", derived using first principles, as one metric the community can use to focus
%attention, in turn making the entire decision-making process dynamic.
%We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
%councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
%improve the legitimacy and efficiency of their decision-making process.
\todo[inline]{TODO}
\end{abstract}

%\category{H.5.m.}{Information Interfaces and Presentation
%  (e.g. HCI)}{Miscellaneous} \category{See
%  \url{http://acm.org/about/class/1998/} for the full list of ACM
%  classifiers. This section is required.}{}{}
%
%\keywords{\plainkeywords}

@@

Traditionally, we can have a stance that everyone had an opportunity to vote and express
their preference, so whatever are results based those who voted are legitimate results.
Research shows, though, that there are many reasons why people do not vote and that there
are many issues related to how elections are organized, i.e., are they held on working
or work-free day.

@@

@@

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented an online application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 5 months.
\item \todo[inline]{We did this analysis and found these results. Legitimacy. Efficiency. Attention.}
\end{itemize}

\todo[inline]{Structure of the paper.}

@@

\section{Related Work}

@@

\section{Decision Credibility}
\label{sec:decision-credibility}
A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.  Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.

@@

@@

\section{Future Work}

\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
Is there a game theoretical approach to it?}

\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
understand what exactly this is and how the math works?}

\section{Conclusion}

>>>>

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Improving legitimacy and efficiency of decisions in democratic communities}
%\def\plainauthor{First Author, Second Author, Third Author,
%  Fourth Author, Fifth Author, Sixth Author}
\def\emptyauthor{}
\def\plainkeywords{Decision making; voting; democracy; quorum; legitimacy; statistical quorum.}
%\def\plaingeneralterms{Documentation, Standardization}

@@

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
maximize productivity?
We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to focus
attention, in turn making the entire decision-making process dynamic.
We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
improve the legitimacy and efficiency of their decision-making process.
\end{abstract}

\category{H.5.3.}{Information Interfaces and Presentation
 (e.g. HCI)}{Computer-supported cooperative work}

\keywords{\plainkeywords}

@@

Traditionally, we can have a stance that everyone had an opportunity to vote and express
their preference, so whatever are results based on those who voted are legitimate results.
Research shows, though, that there are many reasons why people do not vote and that there
are many issues related to how elections are organized~\cite{CANCELA2016264, GIMPEL2003471}.

@@

%Moreover, we can see overall participation by everyone in the population as a
%limited resource, a collective attention, we have at our disposal to make collective decisions.
%Traditionally, a common solution to the low voter turnout is to focus on
%increasing the turnout and through that increasing the collective attention available
%for decision making.
%But increasing collective attention for decision making is at odds with other
%demands for the collective and individual attention.
%But once we have a statistical measure of results, we can utilize it to
%optimize the allocation of the collective attention by directing participation towards
%contentious issues.

@@

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented an online application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases legitimacy of decisions,
improves utilization of collective attention,
and increases number of decisions made while not sacrificing their legitimacy.
\end{itemize}

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do evaluation in
Section~\ref{sec:evaluation}. We discuss results in Section~\ref{sec:discussion} and conclude
in Section~\ref{sec:conclusion}. At the end we have Appendix where we
derive formulas presented in Section~\ref{sec:decision-credibility}.

@@

\section{Related Work}
\label{sec:related-work}

@@

\section{Decision Credibility and Statistical Quorum}
\label{sec:decision-credibility}
A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome of an open vote before the community adopts the outcome as being truly representative all members' positions.
For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose, then we can conclude with high probability that, had all 100 members taken the time to vote, a majority of them would be in support of the proposal.  Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote was credible enough to be taken as representative of the entire community's desire.
On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we cannot yet conclude that the decision to accept the proposal is credible enough for the entire community to officially adopt, despite the significant participation in the vote.
Instead, the suggested course of action would be to continue gathering more votes and/or further discussing the proposal until a credible outcome is established.

@@

\section{Methodology}
\label{sec:methodology}

@@

%\section{Future Work}
%
%\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
%Is there a game theoretical approach to it?}
%
%\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
%understand what exactly this is and how the math works?}

\section{Conclusion}
\label{sec:conclusion}
